---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
dim(counts)
head(counts)
```

look at metadata
```{r}
head(metadata)
```

```{r}
colnames(counts)
```

```{r}
metadata$id
```

check for correspondence of metadata and countdata
```{r}
colnames(counts[-1])
```

```{r}
all(colnames(counts)[-1] ==metadata$id )
```


```{r}
#View(metadata)
```

trying to find the mean for the control vs tx
1.find which columns in **counts** correspond to the control conditions (no drugs)
2.calculate the mean count value for each gene. 
```{r}
control.inds <- metadata$dex == "control"
control.ids <- metadata$id[control.inds]

head(counts[,control.ids])
```

calculate the mean counts value for all genes
```{r}
control.mean <- rowSums(counts[,control.ids])/ncol(counts[,control.ids])
names(control.mean) <- counts$ensgene
head(control.mean)
```

Now do the same thing for the drug tx samples (i.e. **treated** columns)
```{r}
treated.inds <- metadata$dex == "treated"
treated.ids <- metadata$id[treated.inds]

head(counts[,treated.ids])
```
```{r}
treated.mean <- rowSums(counts[,treated.ids])/ncol(counts[,treated.ids])
names(treated.mean) <- counts$ensgene
```
 
 combine!
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

```{r}
colSums(meancounts)
```

create a scatterplot of control vs tx mean
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```

```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log = "xy")
```

the difference between control and mean is in the log2fc column
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

remove genes with zero expression from meancounts to produce **mycounts**
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind = TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
```{r}
sum(up.ind)
```
There are `sum(up.ind)` up regulated genes and `sum(down.ind)` down regulated genes.
```{r}
head(mycounts[up.ind,])
```

Adding annotation data
use the merge() function to add annotation 
```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

```{r}
mycounts.annotated <- merge(x=mycounts, y=anno, by.x ="row.names", by.y = "ensgene")
head(mycounts.annotated)
```


##DESeq2
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results (dds)
head(res)
```

```{r}
res01 <- results(dds,alpha=0.05)
summary(res01)
```

```{r}
mycols <- rep("gray", nrow(res01))
mycols[abs(res01$log2FoldChange)>2] <- "blue"
plot(res01$log2FoldChange, -log(res01$padj), col=mycols)
```

