---
title: "Class 14: Genome Informatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Asthma SNPS in MXL population

Read file from Ensemble
http://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:41279667-41279771;v=rs8067378;vdb=variation;vf=362656008#373531_tablePanel

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

How many GIG genotypes are in this set?
```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
sum(table(mxl$Genotype..forward.strand.))
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl)*100
```

## RNA-seq
Fastq quality scores
Q. Does the first sequence have a good quality?
```{r}
q <- "DDDDCDEDCDDDDBBDDDCC@"
install.packages("seqinr")
install.packages("gtools")
##phred <-asc(s2c(q)) - 33
##phred
```


##Population scale analysis
Does the genotype for rs9067378 influence expression of important asthma linked genes we found?
```{r}
read.csv("rs8067378_ENSG00000172057.6.txt")
```

```{r}
geno <- read.table("rs8067378_ENSG00000172057.6.txt")
head(geno)
```

How many patients are G/G
```{r}
table(geno$geno)
```

get list of trues and falses
```{r}
inds <- geno$geno == "G/G"
summary(geno$exp[inds])
```

```{r}
inds <- geno$geno == "A/G"
summary(geno$exp[inds])
```

```{r}
inds <- geno$geno == "A/A"
summary(geno$exp[inds])
```

Some nice figure that shows the distribution of these expression values?
```{r}
boxplot(exp ~ geno, geno)
```

```{r}
install.packages("contrib.url")
```

