---
title: "Class11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1. The PDB

Q1.Determine the percentage of structures solved by X-Ray and Electron Microscopy.  From the website what proportion of structures are protein?

```{r}
stats <- read.csv("Data Export Summary.csv")
stats
```

Percentage of structures solved by X-ray
```{r}
x <- sum(stats[1,2:5])
y <- sum(stats[,6])
x/y
```

```{r}
percent.by.method <- stats$Total/sum(stats$Total)*100
names(percent.by.method) <- rownames(stats)
percent.by.method
```

Percentage of structures solved by proteins
```{r}
sum(stats$Proteins)/sum(stats$Total)*100
```

##Section 3
```{r}
installed.packages("bio3d")
```

First we have to load our package we want to use
```{r}
library(bio3d)
```
How wany residues? 198
```{r}
pdb <- read.pdb("1hsg")
pdb
```

```{r}
pdb$atom
```

3 letter codes
```{r}
pdb$atom$resid
```

```{r}
aa321(pdb$atom$resid)
```


```{r}
#Read pdb
pdb
```

We want to select out the protein and drug only parts of these molecular PDB files.
```{r}
#Do atom selection for a protein only file
prot.inds <- atom.select(pdb,'protein')
prot.inds
lig.inds <- atom.select(pdb, 'ligand')
lig.inds
```

protein
```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file="protein.pdb")
```

the drug/ligand
```{r}
lig.pdb <- trim.pdb(pdb, lig.inds)
write.pdb(lig.pdb, file="lig.pdb")
```

```{r}
library(bio3d.view)
view(pdb, "overview")
```

